---
title: "05_model_validation"
author: "Loïc Thurre"
date: "2024-09-29"
output:
  pdf_document: default
header-includes: \usepackage{caption}
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output_files") })
---

\captionsetup[table]{labelformat=empty}


```{r setup, include = FALSE}
  knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

  library(dplyr)  
  library(ggplot2)
  library(ggpubr)
  library(stringr)
  library(tidyr)


      # defining plotting theme. 
  MS.theme <- theme(
          axis.text = element_text(size = 15),
          axis.title = element_text(size = 18),
          legend.title = element_text(face = "bold", size = 18),
          legend.text = element_text(size = 17),
          )
  
  pdf.theme <- theme(
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14),
          legend.title = element_text(face = "bold", size = 14),
          legend.text = element_text(size = 13),
  )
  
 
```

# 0. Data Importation

```{r data importation}
  predictions <- read.csv(file = "../input_files/predictions.csv")
  predictions[] <- lapply(predictions, function(x) gsub("\u00A0", "", x))  # Remove non-breaking spaces
  predictions[] <- lapply(predictions, as.numeric) # transform to numeric
  
  
  best_predictors <- read.csv(file = "../input_files/best_predictors.csv", header = FALSE)
  
  predictions <- cbind(predictions, best_predictors[,1])
  colnames(predictions) = c("Observed", "Predicted", "Field")
  predictions$Field <- gsub("\u00A0", "", predictions$Field)
  
  # add BRI data to the dataframe
  soil.bri <- read.csv(file = "../input_files/soil_bri.csv") %>% dplyr::select(c(X, BRI))
  colnames(soil.bri)[1] = "Field"
  
  predictions <- merge(soil.bri, predictions, by = "Field")
  
  load(file = "../input_files/combined_data.rda")
```
# 1. DENOVO leave one out validation 
We computed the LOO validation on the whole approach, the following steps were executed on the N-1 dataset
  1. Splitting between train (N-1) and test set (Nth)
  2. Feature selection 
  3. Best physicochemical model
  4. Best soil fungi model 
  5. Best combined model 
  6. Predictions and comparison with the Nth test site. 
  
This script was executed on the sciCORE and is available in dataset S5 "loo_validation.R". 
We report here the results and produce the corresponding graphs. 

## 1.1 Predictors stability
Selection frequency of predictors during DENOVO LOO selection. 
Counting the number of time a predictor is selected by our approach to compare with what was previously found in dataset S4. 
```{r predictor frequence, echo = FALSE}
  best_predictors[] <- lapply(best_predictors, function(x) gsub("\u00A0", "", x))
  best_predictors$V2 <- str_trim(best_predictors$V2)
  best_predictors$V2 <- str_split(best_predictors$V2, "\\+")
  colnames(best_predictors) = c("Field", "Predictor")
  best_predictors <- unnest(best_predictors, Predictor)
  
  best_predictors$Predictor <- factor(best_predictors$Predictor, unique(best_predictors$Predictor))
  
  predictor_count <- best_predictors %>% count(Predictor) %>%
    arrange(-n)
```

## 1.1 Figure S6A

```{r frequency of predictors}
  relative.predictors <- predictor_count %>% mutate(n = n / nrow(predictions) * 100)

  final.model.pred <- c("WHC", "Corg", "OTU576", "OTU110", "OTU2203", "OTU25")

  fig.6A <- ggplot(relative.predictors, aes(x = reorder(Predictor, n), y = n)) +
    geom_segment(aes(xend=Predictor, 
                     yend=0, 
                     linetype = ifelse(relative.predictors$Predictor %in% final.model.pred, "dashed","solid"))) +
    ylab("Selection frequency (%)")+ xlab("Predictor") +
    geom_point( size=2, color="black") +
    theme_bw() +
    MS.theme +
    theme(
      legend.position = "none",
      plot.margin = unit(c(0,1,0,0), "cm")
    ) +
    coord_flip() 
  
  ggsave(file = "../output_files/figure_S6A.png", plot = fig.6A, width = 6.5, height = 8)
```

```{r figure 6A for pdf, fig.height=8, fig.width = 5}
  fig.6A.pdf <- ggplot(relative.predictors, aes(x = reorder(Predictor, n), y = n)) +
    geom_segment(aes(xend=Predictor, 
                     yend=0, 
                     linetype = ifelse(relative.predictors$Predictor %in% final.model.pred, "dashed","solid"))) +
    ylab("Selection frequency (%)")+ xlab("Predictor") +
    geom_point( size=2, color="black") +
    theme_bw() +
    pdf.theme +
    theme(
      legend.position = "none",
      plot.margin = unit(c(0,1,0,0), "cm")
    ) +
    coord_flip() 

  plot(fig.6A.pdf)
```

## 1.2 Figure S6C
Results of predictions with DENOVO parameters after leave one out validation

```{r Figure 6C}
  fig.6C <- ggscatter(predictions, x = "Observed", y = "Predicted", label = "Field",
            color = "BRI",
            cor.method = "pearson",
            cor.coef = TRUE,
            cor.coef.size = 5,
            point = FALSE,
            font.label = 14,
            repel = T) +
            MS.theme +
  scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
  theme(
    legend.position = 'right'
  )

  ggsave("../output_files/figure_S6C.png", plot = fig.6C, width = 6, height = 4.5, dpi = 300)
```


```{r figure 6C for pdf, fig.width=5, fig.height=4}
 # color this labels with BRI !
  fig.6C.pdf <- ggscatter(predictions, x = "Observed", y = "Predicted", label = "Field",
            color = "BRI",
            cor.method = "pearson",
            cor.coef = TRUE,
            cor.coef.size = 5,
            point = FALSE,
            font.label = 14,
            repel = T) +
            pdf.theme +
  scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
  theme(
    legend.position = 'right'
  )

  plot(fig.6C.pdf)
```

# 2. LOO with selected predictors
Leave one out validation on a model using the previously selected predictors from the combined model: 
WHC, Corg, OTU576, OTU2203, OTU110, OTU25

```{r modelling with selected predictors}
  sel.pred <- data.frame(matrix(ncol = 2))

  for(i in 1:nrow(combined_data)) {
    # split data in training and test sets
    trainData <- combined_data[-i,]
    testData <- combined_data[i,]
    
    # compute model 
    temp.model <- lm(BRI ~ WHC + OTU576 + Corg + OTU2203 + OTU110 + OTU25, data = trainData)
    
    # make prediction
    pred <- predict(temp.model, newdata = testData)
    
    # store prediction in df
    sel.pred[i,] = c(testData$BRI, pred)
  }

  colnames(sel.pred) = c("Observed", "Predicted")
  
  # add field
  sel.pred <- sel.pred %>% mutate(Field = rownames(combined_data))
  
```



## 2.1 Figure S5B
Results of predictions after LOO validation with selected predictors. 

```{r figure 5B}
    # color this labels with BRI !
  fig.5b <- ggscatter(sel.pred, x = "Observed", y = "Predicted", label = "Field",
            color = "Observed",
            cor.method = "pearson",
            cor.coef = TRUE,
            cor.coef.size = 5,
            point = FALSE,
            font.label = 14,
            repel = T) +
            MS.theme +
  scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
  theme(
    legend.position = 'right'
  )

  ggsave("../output_files/figure_S6B.png", plot = fig.5b, width = 6, height = 4.5, dpi = 300)
```


```{r figure 5B for PDF, fig.width=5, fig.height=4}
    # color this labels with BRI !
  fig.5b.pdf <- ggscatter(sel.pred, x = "Observed", y = "Predicted", label = "Field",
            color = "Observed",
            cor.method = "pearson",
            cor.coef = TRUE,
            cor.coef.size = 5,
            cor.coef.coord = c(-40, 10),
            point = FALSE,
            font.label = 14,
            repel = T) +
            pdf.theme +
  scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
  theme(
    legend.position = 'right'
  )

  plot(fig.5b.pdf)
```
  




