---
title: "04_combined_model"
author: "Loïc Thurre"
date: "2024-09-19"
output:
  pdf_document: default
header-includes: \usepackage{caption}
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output_files") })
---

\captionsetup[table]{labelformat=empty}


```{r setup, include = FALSE}
  knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

  library(glmulti)
  library(ggplot2)
  library(dplyr)
  library(phyloseq)
  library(tibble)
  library(olsrr)
  library(relaimpo)
  library(vegan)

    # defining plotting theme. 
  MS.theme <- theme(
          axis.text = element_text(size = 15),
          axis.title = element_text(size = 18),
          legend.title = element_text(face = "bold", size = 18),
          legend.text = element_text(size = 17),
          )
  
  pdf.theme <- theme(
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14),
          legend.title = element_text(face = "bold", size = 14),
          legend.text = element_text(size = 13),
  )
```

We finally explored to which degree a combination of physicochemical and fungal soil parameters could improve the modelling of the observed variability in BRI

# 0. Data importation
```{r data importation}
  # loading phyloseq objects containing soil data + otus count tables and design files
  load("../input_files/mi13.rds")
  load("../input_files/mi13_norm.rds")
  
  # extract soil variables and BRI 
  soil_BRI <- data.frame(sample_data(mi13.norm))
  
  # extract otus data
  otus_BRI <- data.frame(otu_table(mi13.norm))
  
  # join for combined modeling 
  combined_data <- merge(soil_BRI, otus_BRI, by = 0) %>% tibble::column_to_rownames("Row.names")
  
  # save for glmulti on the sciCORE
  save(combined_data, file = "../input_files/combined_data.rda")
```


# 1. Exhaustive screening of the best comined model
Again, we screened for the best OLS model based on previsouly selected otus and soil parameters
```{r}
  # results from the sciCORE
  load("../input_files/combined_glm.rda")

  summary(combined.glm@objects[[1]])
  combined.model <- lm(as.formula(summary(combined.glm)$bestmodel), data = combined_data) # running the model with the best coefficients previously found
```


## 1.1 Normality assumptions
Assessing the residuals normality assumptions the produced model. 
The best model is not fully respecting the residuals normality assumptions. 
These values are aggregated into the Table S5 for each model. 
```{r normality assumptions}
  ols_test_normality(combined.model)
```

## 1.2 Predictors relative importance
```{r relative importance}
   combined.relimp <- calc.relimp(combined.model)
   combined.relimp <- data.frame(combined.relimp@lmg)
   colnames(combined.relimp) = "lmg"
```

## 1.3 Table S10
Summary of the model coefficients and relative importance.
```{r table S10}
  # merge tables so we see the coefficient and relative importance
  combined.coefs <- data.frame(coef(combined.model))
  table.s10 <- merge(combined.coefs, combined.relimp, by = 0, all = TRUE)
  table.s10 <-  table.s10 %>% arrange(lmg)
  colnames(table.s10) = c("Parameter", "Coefficient", "Relative importance")
  
  knitr::kable(table.s10, caption = "Table S10", label = NA)
```


## 1.1 Figure 5A
```{r figure 5A}
   # preparing dataframe for plotting
    combined.relimp <- combined.relimp %>% tibble::rownames_to_column("Parameter") %>% 
    mutate(type = ifelse(grepl("OTU", Parameter), "Fungi", "Soil")) %>%
    mutate(model = "Combined")

  combined.relimp$lmg <- as.numeric(combined.relimp$lmg)
  
  # adding unexplained variance value
  combined.relimp <- rbind(combined.relimp, c("Unexplained", 1-sum(combined.relimp$lmg), "-", "Combined")) 
  combined.relimp$lmg <- as.numeric(combined.relimp$lmg)
  
  combined.relimp <- combined.relimp %>% group_by(type) %>% arrange(-lmg, .by_group = TRUE)
  combined.relimp$Parameter <- factor(combined.relimp$Parameter, levels = combined.relimp$Parameter)
  
  figure.5A <- ggplot(combined.relimp, aes(x = model, y = round(lmg * 100, 2) , fill = Parameter)) +
    geom_col(color = "black", width = 0.25) +
    scale_fill_manual(values = c("white","#0F3045", "#4086AD","#236F9A", "#5F9BBC", "#E5D8D3", "#AE927E")) +
    ylab("% Variance explained") +
    MS.theme +
    theme_void() +
    theme(
      legend.position = "none"
    ) 
  
  
  ggsave(figure.5A, file = "../output_files/figure_5A.png", width = 8, height = 7, dpi = 300)
```


```{r figure 5A for PDF, fig.width = 5, fig.height = 4, fig.align='center'}
  figure.5A.pdf <- ggplot(combined.relimp, aes(x = model, y = round(lmg * 100, 2) , fill = Parameter)) +
    geom_col(color = "black", width = 0.25) +
    scale_fill_manual(values = c("white","#0F3045", "#4086AD","#236F9A", "#5F9BBC", "#E5D8D3", "#AE927E")) +
    ylab("% Variance explained") +
    pdf.theme +
    theme_void() +
    theme(
      legend.position = "none",
    ) 

  plot(figure.5A.pdf)
```



# 2. Redundancy analysis
We used redundancy analysis (RDA) to visualise the relationship between the parameters included in this model and BRI
## 2.1 RDA
```{r redundancy analysis}
  mi13.sqrt <- mi13.norm
  otu_table(mi13.sqrt) <- sqrt(otu_table(mi13.norm)) # sqrt transformation of OTU data.
  
  otus <- data.frame(otu_table(mi13.sqrt))
  env <- data.frame(sample_data(mi13.norm))
  
  rda <- vegan::rda(otus ~ Humus+Clay+Sand+pH+P_H2O+K_H2O+Mg_H2O+Ca_H2O+P_AAE+K_AAE+Mg_AAE+Ca_AAE+Cu+Fe+Mn+B+WHC+Corg+Nmin+CEC, data = env)
  species_scores <- scores(rda, display = "species")
  sites_scores <- scores(rda, display = "sites")
  soil_effects <- data.frame(scores(rda, display = "bp", scaling = 2))
  species_to_highlight <- c("OTU576", "OTU110", "OTU2203", "OTU25")
  soil_to_highlight <- c("WHC", "Corg")
  
  summary(rda)
```

## 2.2 Figure 5B
 We report the combined relationships of chemistry and fungi for all soils in the RDA 
```{r figure 5B}
  figure.5B <-  ggplot() +
      geom_text(data =species_scores, size = 6, angle = 30,aes(x = RDA1*5, y = RDA2*5,
                                    label = ifelse(rownames(species_scores) %in% species_to_highlight, rownames(species_scores), NA))) +
      geom_text(data =sites_scores,  size = 7, aes(x = RDA1, y = RDA2, label = rownames(sites_scores),color = soil_BRI$BRI)) +
        scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
      geom_segment(data = soil_effects, aes(xend = RDA1*2, yend = RDA2*2, x = 0, y = 0,
                   alpha = ifelse(rownames(soil_effects) %in% soil_to_highlight, TRUE, FALSE)),
                   arrow = arrow()) +
      geom_text(data = soil_effects, size = 7, 
                aes(x = RDA1 *2.25, y = RDA2 *2.25, 
                    alpha = ifelse(rownames(soil_effects) %in% soil_to_highlight, TRUE, FALSE), 
                                                   label = rownames(soil_effects))) +
      xlab("RDA1 (9.67 %)") + ylab("RDA2 (8.4 %)") +
      scale_alpha_manual(values = c(0, 1), guide = 'none') +
      theme_bw() +
      MS.theme

  ggsave(figure.5B, file = "../output_files/figure_5B.png", width = 8, height = 7, dpi = 300)
```


```{r figure 5B for pdf, fig.width = 5, fig.height = 4, fig.align='center'}
   figure.5B.pdf <-  ggplot() +
      geom_text(data =species_scores, size = 4, angle = 30,aes(x = RDA1*5, y = RDA2*5,
                                    label = ifelse(rownames(species_scores) %in% species_to_highlight, rownames(species_scores), NA))) +
      geom_text(data =sites_scores,  size = 5, aes(x = RDA1, y = RDA2, label = rownames(sites_scores),color = soil_BRI$BRI)) +
        scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
      geom_segment(data = soil_effects, aes(xend = RDA1*2, yend = RDA2*2, x = 0, y = 0,
                   alpha = ifelse(rownames(soil_effects) %in% soil_to_highlight, TRUE, FALSE)),
                   arrow = arrow()) +
      geom_text(data = soil_effects, size = 5, 
                aes(x = RDA1 *2.25, y = RDA2 *2.25, 
                    alpha = ifelse(rownames(soil_effects) %in% soil_to_highlight, TRUE, FALSE), 
                                                   label = rownames(soil_effects))) +
      xlab("RDA1 (9.67 %)") + ylab("RDA2 (8.4 %)") +
      scale_alpha_manual(values = c(0, 1), guide = 'none') +
      theme_bw() +
      pdf.theme

  plot(figure.5B.pdf)
```



















