---
title: "02_Soil"
author: "Loïc Thurre"
date: "2024-09-17"
output:
  pdf_document: default
header-includes: \usepackage{caption}
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output_files") })
---

\captionsetup[table]{labelformat=empty}


```{r setup, include = FALSE}
  knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

  library(ggplot2)
  library(dplyr)
  library(glmulti)
  library(ggfortify)
  library(tibble)
  library(openxlsx)
  library(psych)
  library(svglite)
  library(reshape2)
  library(relaimpo)
  library(olsrr)
  library(purrr)

  # defining plotting theme. 
  MS.theme <- theme(
          axis.text = element_text(size = 15),
          axis.title = element_text(size = 18),
          legend.title = element_text(face = "bold", size = 18),
          legend.text = element_text(size = 17),
          )
  
  pdf.theme <- theme(
          axis.text = element_text(size = 12),
          axis.title = element_text(size = 14),
          legend.title = element_text(face = "bold", size = 14),
          legend.text = element_text(size = 13),
  )
```

# 0. Data importation
```{r data importation}
    # Importing soil parameters
    SOIL <- read.xlsx(xlsxFile = "../input_files/soil_parameters.xlsx", 
                     sheet = 1)  %>% column_to_rownames("soil_ID")

    # importing unfertilised BRI data from S1
    BRI.unfert <- read.csv(file = "../input_files/BRI_unfertilised.csv") %>%
      mutate(field = gsub("soil_ID", "", new_field)) %>% tibble::column_to_rownames("field") %>% dplyr::select(Estimate)
    
    # merging BRI and soil tables
    masterTable <- merge(BRI.unfert, SOIL, by = 0) %>% column_to_rownames("Row.names")
    colnames(masterTable)[1] = "BRI" # renaming Estimate into BRI
    
    write.csv(masterTable, file = "../output_files/soil_bri.csv")
```

# 1. Principal coordinate analysis (PCA)
First, we performed a principal component analysis (PCA) to reveal relationships between the assessed physicochemical parameters across the 25 soils 
# 1.1 Computing PCA
```{r soil PCA}
  soil.pca <- prcomp(masterTable[, -1], scale. = TRUE, center = TRUE)
  summary(soil.pca)
```



## 1.2 Figure 3A
To obtain insights on the relationship of soil parameters and BRI, we coloured the soils in the PCA based on the BRI of wheat growth on these soils
```{r fig.3A}
  fig.3A <- autoplot(soil.pca, data = masterTable, 
           colour = "BRI", 
           loadings = TRUE, 
           loadings.label = TRUE,
           loadings.color = "darkgray",
           loadings.label.color = "darkgray",
           loadings.label.size = 6,
           label = TRUE,
           label.size = 7,
           shape = FALSE) + 
    scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
    theme_bw() +
    MS.theme

    ggsave(fig.3A, file = "../output_files/figure_3A.svg", width = 8, height = 6, dpi = 300) # saving as SVG for subsequent modification with illustrator
```

```{r fig.3A for pdf, fig.width = 7, fig.height = 4}
  fig.3A.pdf <- autoplot(soil.pca, data = masterTable, 
           colour = "BRI", 
           loadings = TRUE, 
           loadings.label = TRUE,
           loadings.color = "darkgray",
           loadings.label.color = "darkgray",
           loadings.label.size = 4,
           label = TRUE,
           label.size = 5,
           shape = FALSE) + 
    scale_color_gradient( low = "#0f5dc1", high = "#2dfa88", name = "BRI (%)") +
    theme_bw() +
    pdf.theme

    plot(fig.3A.pdf)
```


## 1.3 ANOVA 
Formula: BRI ~ soil_variables
```{r ANOVA BRI ~ soil_variables}
  shapiro.test(masterTable$BRI)
  
  soil.anova <- aov(BRI ~ Humus+Clay+Silt+Sand+pH+P_H2O+K_H2O+Mg_H2O+Ca_H2O+Nitrate+P_AAE+K_AAE+Mg_AAE+Ca_AAE+Cu+Fe+Mn+B+WHC+Corg+SOM+Nmin+CEC, masterTable)
```

## Table S3? 
```{r table S3}
  knitr::kable(anova(soil.anova), caption = "Table S3", label = NA)
```


# 2. Pairwise correlation BRI ~ soil_variable
Next, we assessed pairwise correlations between the physicochemical soil parameters and BRI. 
```{r co-correlation between variables}
  bri.soil.corr <- corr.test(masterTable$BRI, masterTable, use = "pairwise", method = "spearman", adjust = "bonferroni")
```

## Table S4
```{r table S4}
  table.s4 <- cbind(t(bri.soil.corr$r), t(bri.soil.corr$p), t(bri.soil.corr$p.adj))
  colnames(table.s4) = c("Spearman coefficient", "P-value", "Adjusted P-value")
  knitr::kable(table.s4[-1,], caption = "Table S4", label = NA)
```


# 2. Removal of co-correlated predictors
We reduced the sets of parameters (removal of co-correlated variables). We exclude the co-correlated predictors (|r| > 0.8).
```{r co-correlation soil predictors, echo = FALSE}
  soil.corr <- corr.test(SOIL, method = "spearman", adjust = "bonferroni")
  corr.matrix <- soil.corr$r
  corr.matrix[upper.tri(corr.matrix, diag = TRUE)] = NA
  temp <- melt(corr.matrix, na.rm = TRUE)

  soil.corr <- cbind(temp, soil.corr$p.adj)
  excluded <- c("Silt", "Nitrate", "SOM")
  
  soil_selected <- SOIL %>% dplyr::select(!excluded)
```

## 2.1 Figure S5
```{r figure S5}
  figure.s5 <- ggplot(soil.corr, aes( x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    theme_bw() +
    scale_fill_gradient2(low = "red", mid="white", high="blue") +
    MS.theme +
    theme(
      axis.text.x = element_text(angle = 90,
                                 face = ifelse(soil.corr$Var1 %in% excluded, "plain", "bold")),
      axis.title = element_blank(),
      legend.title = element_blank()
    )
  ggsave(file = "../output_files/figure_s5.png",
       plot = figure.s5,
       dpi = 300,
       width = 10,
       height = 9)
```

```{r figure S5 for PDF}
  figure.s5.pdf <- ggplot(soil.corr, aes( x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    theme_bw() +
    scale_fill_gradient2(low = "red", mid="white", high="blue") +
    pdf.theme +
    theme(
      axis.text.x = element_text(angle = 90,
                                 face = ifelse(soil.corr$Var1 %in% excluded, "plain", "bold")),
      axis.title = element_blank(),
      legend.title = element_blank()
    )

  plot(figure.s5.pdf)
```

# 3. Exhaustive screening of candidate models with glmulti()
To find the best model we proceed to an exhaustive screening of all the possible models BRI ~ soil_variables with the AICc criterion. 
```{r find best soil model}
  BRI.soil <- merge(BRI.unfert, soil_selected, by = 0) %>% column_to_rownames("Row.names")
  colnames(BRI.soil)[1] = "BRI"
  lm.soil <- lm(BRI ~. , data = BRI.soil)
  
  # This step is ran on the sciCORE
  save(BRI.soil, file = "../output_files/BRI_soil.rda") # to send on the sciCORE
  # soil.glm <- glmulti(lm.soil, level = 1, plotty = F, trace = F, crit = aicc)
  
  # We load the glmulti object from the sciCORE. 
  load("../input_files/soil_glm.rda")
  summary(soil.glm@objects[[1]])
  
  
  soil.model <- lm(as.formula(summary(soil.glm)$bestmodel), data = BRI.soil) # running the model with the best coefficients previously found
```

## 3.1 Normality assumptions of the model
Assessing the residuals normality assumptions the produced model. 
These values are aggregated into the Table S5 for each model. 

```{r soil model normality assumptions}
  norm.soil <- ols_test_normality(soil.model)
  print(norm.soil)
```


## 3.2 Relative importance of best model 
The relative importance of each predictor in the previously selected model is computed in this step. These values are then used to complete the figure 3B in illustrator. 
```{r soil model predictor relative importance}
  soil.relimp <- calc.relimp(soil.model)
  soil.relimp <- data.frame(soil.relimp@lmg)
  colnames(soil.relimp) = "lmg"
```



## 3.3 Table S6
Summary of the model coefficient and relative importance 
```{r Table S6}
  # merge tables so we see the coefficient and relative importance
  soil.coefs <- data.frame(coef(soil.model))
  table.s6 <- merge(soil.coefs, soil.relimp, by = 0, all = TRUE)
  table.s6 <-  table.s6 %>% arrange(lmg)
  colnames(table.s6) = c("Parameter", "Coefficient", "Relative importance")
  
  knitr::kable(table.s6, caption = "Table S6", label = NA)
```


## 3.4 Figure 3B
```{r Figure 3B}
  # preparing dataframe for plotting
    v2.soil.relimp <- soil.relimp %>% tibble::rownames_to_column("Parameter") %>% 
    mutate(type = ifelse(grepl("OTU", Parameter), "Fungi", "Soil")) %>%
    mutate(model = "Soil")

  v2.soil.relimp$lmg <- as.numeric(v2.soil.relimp$lmg)
  # adding unexplained variance value
  v2.soil.relimp <- rbind(v2.soil.relimp, c("Unexplained", 1-sum(v2.soil.relimp$lmg), "-", "Soil")) 
  v2.soil.relimp$lmg <- as.numeric(v2.soil.relimp$lmg)
  
  v2.soil.relimp <- v2.soil.relimp %>% group_by(type) %>% 
    arrange(-lmg, .by_group = TRUE) %>% 
    map_df(rev) %>% 
    mutate(added_lmg = cumsum(lmg)) %>% 
    map_df(rev)
  v2.soil.relimp$Parameter <- factor(v2.soil.relimp$Parameter, levels = v2.soil.relimp$Parameter)
  
  figure.3b <- ggplot(v2.soil.relimp, aes(x = model, y = round(lmg * 100, 2) , fill = Parameter)) +
    geom_col(color = "black", width = 0.25) +
    scale_fill_manual(values = c("white", "#E5D8D3", "#AE927E", "#593112", "#30190C")) +
    ylab("% Variance explained") +
    theme_void() +
      theme(legend.position = "none")
  
  ggsave(filename = "../output_files/figure_3B.png", figure.3b)
```

```{r figure 3B for pdf, fig.width = 5, fig.height = 4, fig.align='center'}
  figure.3B.pdf <- ggplot(v2.soil.relimp, aes(x = model, y = round(lmg * 100, 2) , fill = Parameter)) +
    geom_col(color = "black", width = 0.25) +
    scale_fill_manual(values = c("white", "#E5D8D3", "#AE927E", "#593112", "#30190C")) +
    ylab("% Variance explained") +
    pdf.theme +
    theme_void() +
      theme(legend.position = "none") 

  plot(figure.3B.pdf)
```


